Assessment of zonally symmetric and asymmetric components of the Southern Annular Mode using a novel approach

### 2021-04-30 analysis/paper/asymsam.Rmd
Aclara mejor diferencias entre DJF y otros.
#17

By performing the EOF analysis using data for all months we are assuming that the zonally asymmetric structure of the SAM is the same at all seasons.
The latter was assessed by computing geopotential height zonal anomalies by projecting the first EOF of each season independently.
The following seasons were considered -- December to February (DJF), March to May (MAM), June to August (JJA) and September to November (SON).
Results are very similar to each other in the troposphere (Figure \@ref(fig:A5), row 2) and show spatial correlations between 0.65 (DJF with JJA) and 0.9 (MAM with SON).
In the stratosphere (Figure \@ref(fig:A5), row 1), patterns are similar for all seasons except DJF, when the wave-1 zonal anomalies are rotated 90\degree in comparison with the rest of the year.
Spatial correlations in the stratosphere are between -0.24 (DJF with SON) and 0.95 (MAM with JJA).
Therefore, the results confirm that the zonal asymmetric structure of the SAM is very similar throughout most of the year.
The only outlier being DJF, which shows much lower correlations with the other seasons at both levels and the weakest zonal anomalies (Figure \@ref{fig:A3}), which is consistent with @fogt2020.

@@

patterns <- SAM_season[, eof[[1]]$left, by = .(lev, season)] %>% 
  copy() %>% 
  setnames("hgt", "EOF") %>% 
  .[, EOF := EOF/sd(EOF), by = .(season, lev)] %>% 
  hgt[., on = c("time", "lev")] %>%
  .[, FitLm(hgt_a, EOF), by = .(lon, lat, lev, season)] %>%
  rm_intercept() %>% 
  .[season == "DJF", estimate := -estimate] %>% 
  .[, estimate_a := Anomaly(estimate), by = .(lat, lev, season)]

>>>>

By performing the EOF analysis using data for all months we are assuming that the zonally asymmetric structure of the SAM is the same at all seasons.
The latter was assessed by computing geopotential height zonal anomalies by projecting the first EOF of each season independently.
The following seasons were considered -- December to February (DJF), March to May (MAM), June to August (JJA) and September to November (SON).
Results are very similar to each other in the troposphere (Figure \@ref(fig:A5), row 2) and show spatial correlations between 0.65 (DJF with JJA) and 0.9 (MAM with SON).
In the stratosphere (Figure \@ref(fig:A5), row 1), patterns are similar for all seasons except DJF, when the wave-1 zonal anomalies are rotated 90\degree in comparison with the rest of the year.
Spatial correlations in the stratosphere are between -0.24 (DJF with SON) and 0.95 (MAM with JJA).
Therefore, the results confirm that the zonal asymmetric structure of the SAM is very similar throughout most of the year.
The outlier being DJF, which shows much lower correlations with the other seasons at both levels and the weakest zonal anomalies (Figure \@ref{fig:A3}), which is consistent with @fogt2020.
Therefore, even though we do our analysis for all months, we would expect for it to represent the other trimesters better.

@@

patterns <- SAM_season[, eof[[1]]$left, by = .(lev, season)] %>% 
  copy() %>% 
  setnames("hgt", "EOF") %>% 
  .[, EOF := EOF/sd(EOF), by = .(season, lev)] %>% 
  hgt[., on = c("time", "lev")] %>%
  .[, FitLm(hgt_a, EOF), by = .(lon, lat, lev, season)] %>%
  rm_intercept() %>% 
  .[season == "DJF", estimate := -estimate] %>% 
  .[, estimate_a := Anomaly(estimate), by = .(lat, lev, season)] %>% 
  .[season == "DJF" & lev == 700, estimate_a := -estimate_a]
